A Reproducible Workflow with RMarkdown, Git, Make & Docker

### 2019-08-20 rr-flow.Rmd
add references

Reproducible Research has become the gold standard for scientific work in recent years. The following tutorial aims to outline a workflow that guarantees not only a reasonable standard of reproducibility but also decreases the burden of maintaining software for the researcher and her colleagues. The authors' believe that integrating tools from software engineering in the process of scientific reporting to enhance reproducibility is crucial and give guidelines for the practitioner to use them. These tools include software for literate programming, version control, dependency tracking, and containerization. If implemented, such workflow enables the researcher to transparently create, dynamicly render and puplicly share reports, which she and her colleages can trust to reproduce later, regardless of operating system or software version.

Literate Programming
: Interwoven text and computer code that together forms a human-readable document.
: Example: One document that includes text that describes a statistical test, code that is used to execute the statistical and the output of the code in the same document.
Version Control
: The management of change of documents/code.
: Example: Changes from the manuscript text upon request by the reviewer are identifiable.

@@

Within the community of users of the R Programming Language for each purpose a specific software solutions is clearly the most popular. For literate programming, `rmarkdown`, for version control `git`, for dependency management `make` and for container `docker`. Each of them serves a meta-scientific valuable goal (reproducibility) and increases the producitivity of the researcher. Each of these software solutions are extremly mighty, so in order to master them fully often years of practice are required. Luckily the minimal valuable subset needed to ensure reproducibility can be learned with little practice. However even if only a part of this workflow is implemented, the

>>>>

Reproducible Research has become the minimal standard for scientific work [@hardwicketome.DataAvailabilityReusability] in recent years. The following tutorial aims to outline a workflow that guarantees not only a reasonable standard of reproducibility but also decreases the burden of maintaining software and code for the researcher and her colleagues. The authors' believe that integrating tools from software engineering in the process of scientific reporting to enhance reproducibility is crucial and give guidelines for the practitioner to use them. These tools include software for literate programming, version control, dependency tracking, and containerization. If implemented, such workflow enables the researcher to transparently create, dynamicly render and puplicly share reports, which she and her colleages can trust to reproduce later, regardless of operating system or software version.

Literate Programming
: Interwoven text and computer code that together forms a human-readable document.
: Example: One document that includes text that describes a statistical test, code that is used to execute the statistical and the output of the code in the same document.

Version Control
: The management of change of documents/code.
: Example: Changes from the manuscript text upon request by the reviewer are identifiable.

@@

While it is widely recognized that these tools enhance reproducibility and there are several guidelines for using them (Literate Programming: @ruleTenSimpleRules2019; Version Control: @barbaHardRoadReproducibility2016; Dependency Management: @askrenUsingMakeReproducible2016; Containerization: @clyburne-sherinComputationalReproducibilityContainers2018) Within the community of users of the R Programming Language for each purpose a specific software solutions is clearly the most popular. For literate programming, `rmarkdown`, for version control `git`, for dependency management `make` and for container `docker`. Each of them serves a meta-scientific valuable goal (reproducibility) and increases the producitivity of the researcher. Each of these software solutions are extremly mighty, so in order to master them fully often years of practice are required. Luckily the minimal valuable subset needed to ensure reproducibility can be learned with little practice.
